### 2014-04-02 paper/tubes.tex
comments on latter half of the paper
first set of glamour shots, paper work, redo of figures on new mm background, redo some figures due to new modeling

\section{A Series of Tubes - A Tool for Pipe Design}

@@

Beyond design, the tubes must actually be printable to be useful.  When geometries contain many overhangs, 3D printers lay down support material that is later removed from the model.  This removal process can be time-intensive and challenging.  The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  We will discuss techniques for fused deposition modeling (FDM) machines like the Makerbot and multi-jetting machines like the Objet.

FDM machines lay down lines of plastic on top of each other: this allows ``stepover'', when a layer of plastic is laid slightly offset from the layer below it, and ``bridging'', when plastic is laid across a gap between two surfaces.  For a layer to fix correctly to its supporting layer, roughly half of the filament needs to overlap.  This means that overhangs without added support material can be approximately $45^{\circ}$, and although cylinders (e.g., pipes) and spheres violate this in some places they are printable support-free.  When support is laid, it is made from the same material as the model in a heatsink-like wave pattern.  It is straightforward to remove from exterior model surfaces with fingers or pliers, but nearly impossible to remove from internal cavities.

On multi-jetting machines, bridging is not possible and stepover potential is much less, because the machine is laying down a liquid rather than a plastic.  Unsupported overhangs can be approximately $14^{\circ}$.

\bjoern{Argue first that we need processes to fabricate these hollow pipes- can do this by removing support material or never printing such material in the first place.}  \bjoern{Is section needs to bee written from a general standpoint first, and then our specific details second.}
The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  The two machines we used to create our example objects were the Objet Connex 260 and the Makerbot Replicator 2.

The Objet is an inkjet-based 3D printer which jets drops of polymer in very fine layers, curing each layer with an ultraviolet lamp.  It offers a wide range of material types, including clear/opaque, flexible/hard, and black/white.  Additionally, it permits the mixing of two different materials in a single print: a mixed material could include 70\% flexible and 30\% hard, for example, to be firm but somewhat flexible.  The flexible materials can be inflated when they are thin: they elongate 200\% before breaking.  We have found $.05$ inches to be a good thickness for inflatable materials.  Although this machine has many extraordinary capabilities, it unfortunately cannot print overhangs of greater than $14^{\circ}$ without support material.  Due to this, support avoidance is difficult; our main tactic for removing support material has been to make the \emph{openings} to pipes as large as possible.  This allows us to flush the material most easily.  The support structure used by this machine is a loose matrix of material that can be removed by hand from the surface, blasted out with a water jet from cavities, or dissolved in lye (the enclosing model material does not dissolve).

The Makerbot Replicator 2 is a fused-deposition modeling (FDM) printer, in which we use exclusively Polylactic Acid (PLA) bioplastics.  More material properties are being explored in PLA: there are now transparent and flexible materials.  They cannot be blended like those in the Objet.  The Makerbot has a fairly coarse resolution compared to the Objet, however it permits overhangs up to $45^{\circ}$ without support and also allows bridging (short spans of material that bridge the gap between two structures with no support underneath).  Although spheres and sideways cylinders have a $>45^{\circ}$ overhang in some places, they are also printable without support.  Due to the FDM process it can even recover from greater overhangs and longer bridges, as the material builds up and can reorganize into a structure \valkyrie{how do I say this?  do I need a photo?  print recovery is pretty awesome-looking.}.  The support used by the Makerbot is made of the same material as the model itself, just printed in a different pattern and attached over as small a surface area as possible to ease removal.  Support inside of a pipe may require pliers, or be infeasible to remove dependent upon geometry.  Meshmixer has a built-in tool for minimizing support for prints; it re-orients the part to keep as much surface area as possible within the 45 degree constraint.  In future work we would like to explore a tool in which a user can select surfaces to prioritize support avoidance for.

Assembly is easiest on both machines when the model does not require support material.  For especially complex geometries, models can be ``cut up'' \bjoern{This negates some of the benefits compared to other manufacturing techniques.} into assembleable pieces that evade support requirements or ease material removal.  We believe that an automatic tool for this task would be highly beneficial, however it is outside the scope of this work.

>>>>

\section{A Series of Tubes - A Tool for Pipe Design}
\bjoern{What kind of visualizations and diagrams can help explain the technical part other than photos and screenshots? The "UIST" pipe routing figure goes in this direction. Can we make similar conceptual figures, flowcharts, etc for the earlier parts?}

@@

Beyond design, the tubes must actually be printable to be useful.  When geometries contain overhangs, as there may be in designed pipes and cavities, 3D printers lay down support material that can be later removed from the model.  This removal process can be time-intensive and challenging.  The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  We will discuss techniques for fused deposition modeling (FDM) machines like the Makerbot and multi-jetting machines like the Objet.

FDM machines lay down lines of plastic on top of each other: this allows ``stepover'', when a layer of plastic is laid slightly offset from the layer below it, and ``bridging'', when plastic is laid across a gap between two surfaces.  For a layer to fix correctly to its supporting layer, roughly half of the filament needs to overlap.  This means that overhangs without added support material can be approximately $45^{\circ}$, and although cylinders (e.g., pipes) and spheres violate this in some places they are printable support-free.  Support is made from the same material as the model, laid in a heatsink-like wave pattern.  It is straightforward to remove from exterior surfaces with fingers or pliers, but nearly impossible to remove from internal cavities.  Meshmixer has a built-in tool for minimizing support for FDM prints; it re-orients the part to keep as much surface area as possible within the 45 degree constraint.  In future work we would like to explore a tool in which a user can select surfaces to prioritize support avoidance for.

On multi-jetting machines, bridging is not possible and stepover potential is much less, because the machine is laying down droplets of liquid rather than more structural plastic.  Unsupported overhangs can be approximately $14^{\circ}$.  The support structure used by this machine is a loose matrix of material that can be removed by hand from the surface, blasted out with a water jet from cavities, or dissolved in lye (the enclosing model material does not dissolve).  Thus, any cavity to which there is some access can be cleared of support.

For especially complex geometries, models can be ``cut up'' into assembleable pieces that evade support requirements or ease material removal.  This is undesirable as it means more time spent in assembly post-print, and additionally a cut pipe may not be fluid-tight any longer.  An automatic tool for this task might be beneficial, however it is outside the scope of this work.
